From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-17 paper.tex
Tentative rewrite of the abstract with the concept of with-me-ness

\title{Real-time Attention Assessment for Social Robots}
\author{Fernando Garcia \qquad Séverin Lemaignan \qquad Alexis Jacq \qquad Pierre Dillenbourg\\Computer-Human Interaction in Learning and Instruction Laboratory (CHILI)\\École Polytechnique Fédérale de Lausanne (EPFL)}

@@

Successfully sustaining face-to-face interaction between humans and robots
mandates fine perception of each other attentional focus. Along this line, we
present a novel technique to assess the visual focus of attention (VFoA) based
on accurate, real-time head pose estimation with monocular camera. The software
is fully open-source and can readily be adopted on a wide range of social
robots. We propose an in-depth evaluation of the technique, using manual video
annotations as ground truth, and show that our system reaches high level of
accuracy in a real-world face-to-face interaction scenario. We finally suggest how
to derive a metric of engagement from the monitoring of the VFoA, allowing the
robot to account and react to the user's changing engagement in real-time.

>>>>

\title{From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction}
\author{Fernando Garcia \qquad Séverin Lemaignan \qquad Alexis Jacq \qquad Pierre Dillenbourg\\Computer-Human Interaction in Learning and Instruction Laboratory (CHILI)\\École Polytechnique Fédérale de Lausanne (EPFL)}

@@

Successfully sustaining face-to-face interaction between humans and robots
mandates fine perception of each other attentional focus. This article presents
a novel open-source technique to assess the visual focus of attention (VFoA)
based on accurate, real-time head pose estimation with a single monocular
camera.  Using manual video annotations as ground truth, we show that our system
reaches satisfactory levels of accuracy in a real-world face-to-face interaction
scenario.

We then propose to introduce the concept of ``with-me-ness'', borrowed from the
field of {\it Computer-Supported Collaborative Learning}, as an
\emph{in-the-moment} measurement of the quality of interaction. We argue that
this new metric, more specific than the broad concept of engagement, may be well
suited for operational, real-time assessment of an on-going interaction.
